Centralized Programming for Distributed Control Planes

### 2016-01-26 paper/propane.tex
make prettier code

The operator configures the network with the abstraction that he or she has centralized control over routing and uses a set of high-level constraints to describe the different routes traffic may (or may not) take and their relative preferences.
In general, \sysname specifications are written modularly via a series
of user declarations.
%These definitions allow users to express the three
%major elements of any \sysname specification:  \emph{prefixes},
%\emph{paths} and \emph{policies}.
For example, to begin specification of the backbone
network, we first express the idea that for routes leading to our customer,
we prefer using $R_1$ over $R_2$ (Policy P3 from Figure~\ref{fig:example1}):
\begin{lstlisting}[mathescape=true]
define ExitCust = $exit(R_1 \gg R_2)$
\end{lstlisting}
The  statement above defines a set of \emph{ranked paths}, which includes
all paths (and only those paths) that \cd{exit}
our network
through either router $R_1$ or $R_2$.  The paths that exit through $R_1$
are preferred ($\gg$) to the paths that exit through $R_2$.

To associate ranked paths with
one or more prefixes, we must define a \sysname \emph{policy}.
Within a policy, statements with the form $\path{x}{p}$
associate the prefixes defined by the predicate $x$ with the set of
ranked paths defined by the path expression $p$.
For instance, in the following code, ranked paths are associated with
the customer prefixes ($C_{pfx}$), the internal prefixes ($I_{pfx}$),
and all other prefixes (\cd{true}).\footnote{Policy statements are processed in
order with earlier policy statements taking precedence over later
policy statements.  Hence, when the prefix predicate \cd{true} follows
statements involving $C_{pfx}$ and $I_{pfx}$, it is interpreted as
$\cd{true} \AND \NOT C_{pfx} \AND \NOT I_{pfx}$.}

\begin{lstlisting}[mathescape=true]
define Routing = {
    $\path{C_{pfx}}{ExitCust \gg end(Cust)}$
    $\path{I_{pfx}}{end(in)}$
    $\path{true}{ExitCust \gg exit(Peer) \gg exit(Prov)}$
}
\end{lstlisting}

Line 2 of the policy above
defines the paths that exit our network to customer prefixes ---
\cd{ExitCust} defines paths through $R_1$ and $R_2$ and in the event
that connections to the customer through both of those routers fail,
a backup route ($end(Cust)$) is defined that admits traffic to go through
any path that ends at the customer.
Line 3 states that traffic for internal prefixes must end in our network, and is otherwise unconstrained.  The special keyword \cd{in} represents any location
inside the user's network whereas the keyword \cd{out} represents any location
outside the user's network.
Line 4 applies to any other traffic, and allows for any routes that leave through a peer with a preference for customers over peers over providers. To summarize our progress so far, the \cd{Routing} policy
implements (P1), (P3) and (P4) from Figure~\ref{fig:example1}

Since, routing still allows transit traffic (e.g., traffic can enter from \textit{Peer} and leave through \textit{Prov}), we can define a new policy that
implements this restriction separately.

\begin{lstlisting}[mathescape=true]
define PP = Peer or Prov
define PPTransit = $enter(PP) \AND exit(PP)$
define CustTransit = $later(Cust) \AND later(Prov)$

define NoTransit = {
    $\path{true}{\neg PPTransit \AND \NOT CustTransit}$
}
\end{lstlisting}

The \textsf{NoTransit} constraint above ensures that requirements (P2) and (P5) are satisfied. In particular, it says that for any prefix, traffic should never both enter and exit the network from \textit{Peer} or \textit{Prov}. Similarly it prevents traffic from ever following paths that leave our network and later passing through both \textit{Prov} and \textit{Cust}.  To implement both \cd{Routing}
and \cd{NoTransit} simultaneously, we simply conjoin them:

\begin{lstlisting}[mathescape=true]
Routing $\AND$  NoTransit
\end{lstlisting}

The lines above capture the entire policy. From it, \sysname will generate per-device import and export filters, local preferences,
MED attributes, and community tags to ensure that the policy is
implemented correctly under all failure cases.

@@

In our data center example,
there are primarily three main concerns:
(1) traffic for the prefix block allocated to each top-of-rack router must be able to reach that router,
(2) local services must not leak outside the datacenter, and
(3) aggregation must be performed on global prefixes to reduce churn
in the network.

@@

\begin{lstlisting}[mathescape=true]
define Ownership = {
    $\path{p_{g1}}{end(A)}$
    $\path{p_{g2}}{end(B)}$
    $\path{p_{l1}}{end(E)}$
    $\path{p_{l2}}{end(F)}$
}
\end{lstlisting}

In English: traffic for prefix $p_{g1}$ is only allowed to follow paths that
end at router A; traffic not matching $p_{g1}$, but which matches $p_{g2}$ must
end at router B; and so on.

@@

\begin{lstlisting}[mathescape=true]
define Routing = {
    $\path{p_{g*}}{any}$
    $\path{p_{l*}}{\NOT enter(out)}$
    $\path{true}{exit(out)}$
}
\end{lstlisting}

The first line states that there is no restriction (\cd{any})
on how traffic must
traverse the network for global prefixes, aside from the default restriction
that traffic must not pass through the user's network and then loop
back on itself. This means traffic for
$p_{g*}$ may be sent either from other routers in the datacenter, or
from external ASes. The second line ensures that traffic for local
prefixes never enters the network from an outside location. This constraint
guarantees that the services remain reachable only to locations
internal to the data center.

As in the backbone example, we can combine these constraints
logically to specify the network-wide policy.
However, in addition to constraints on the shape of paths,
\sysname allows the operator to specify constraints on the BGP control plane.
For instance, a constraint on aggregation is included to ensure that
aggregation for $p_{agg}$ is performed from locations inside the network
to locations outside the network. \todo{Say more about aggregation here.
What prefixes are aggregated in to what blocks?  How is this defined?}

\begin{lstlisting}[mathescape=true]
Ownership $\wedge$ Routing $\wedge$ agg($p_{agg}$, $in \rightarrow out$)
\end{lstlisting}

>>>>

The operator configures the network with the abstraction that he or she has centralized control over routing and uses a set of high-level constraints to describe the different routes traffic may (or may not) take and their relative preferences.
In general, \sysname specifications are written modularly via a series
of user declarations.
%These definitions allow users to express the three
%major elements of any \sysname specification:  \emph{prefixes},
%\emph{paths} and \emph{policies}.
For example, to begin specification of the backbone
network, we first express the idea that for routes leading to our customer,
we prefer using $R_1$ over $R_2$ (Policy P3 from Figure~\ref{fig:example1}):
%\begin{lstlisting}[mathescape=true]
%define ExitCust = $exit(R_1 \gg R_2)$
%\end{lstlisting}
\begin{code}
\Define ExitCust = \Exit(R1 \Prefer R2)
\end{code}
The  statement above defines a set of \emph{ranked paths}, which includes
all paths (and only those paths) that exit
our network
through either router \CD{R1} or \CD{R2}.  The paths that exit through \CD{R1}
are preferred (\Prefer) to the paths that exit through $R_2$.

% DAVE CONVERSIONS
% C_{pfx} ===> PCust
% I_{pfx} ===> PInternal

To associate ranked paths with
one or more prefixes, we must define a \sysname \emph{policy}.
Within a policy, statements with the form $x$\Path$p$
associate the prefixes defined by the predicate $x$ with the set of
ranked paths defined by the path expression $p$.
For instance, in the following code, ranked paths are associated with
the customer prefixes (\CD{PCust}), the internal prefixes (\CD{PInt}),
and all other prefixes (\CD{true}).\footnote{Policy statements are processed in
order with earlier policy statements taking precedence over later
policy statements.  Hence, when the prefix predicate \True follows
statements involving \CD{PCust} and \CD{PInt}, it is interpreted as
\CD{\True{} \AND{} \NOT{}PCust \AND{} \NOT{}PInt}.}

%% \begin{lstlisting}[mathescape=true]
%% define Routing = {
%%     $\path{C_{pfx}}{ExitCust \gg end(Cust)}$
%%     $\path{I_{pfx}}{end(in)}$
%%     $\path{true}{ExitCust \gg exit(Peer) \gg exit(Prov)}$
%% }
%% \end{lstlisting}

\begin{code}
\Define Routing = \{
     PCust \Path ExitCust \Prefer \End(Cust)
     PInt  \Path \End(in)
     \True  \Path ExitCust \Prefer \Exit(Peer) \Prefer \Exit(Prov)
\}
\end{code}
\noindent
Line 2 of the policy above
defines the paths that exit our network to customer prefixes ---
\CD{ExitCust} defines paths through \CD{R1} and \CD{R2} and in the event
that connections to the customer through both of those routers fail,
a backup route (\CD{\End(Cust)}) is defined that admits traffic to go through
any path that ends at the customer.
Line 3 states that traffic for internal prefixes must end in our network, and is otherwise unconstrained.  The special keyword \In{} represents any location
inside the user's network whereas the keyword \Out{} represents any location
outside the user's network.
Line 4 applies to any other traffic, and allows for any routes that leave through a peer with a preference for customers over peers over providers. To summarize our progress so far, the \CD{Routing} policy
implements (P1), (P3) and (P4) from Figure~\ref{fig:example1}.

Since routing allows transit traffic (\emph{i.e.}, traffic can enter from 
\CD{Peer} and leave through \CD{Prov}), we should define a new policy that
implements this restriction separately.

%% \begin{lstlisting}[mathescape=true]
%% define PP = Peer or Prov
%% define PPTransit = $enter(PP) \AND exit(PP)$
%% define CustTransit = $later(Cust) \AND later(Prov)$

%% define NoTransit = {
%%     $\path{true}{\neg PPTransit \AND \NOT CustTransit}$
%% }
%% \end{lstlisting}
\begin{code}
\Define PP = Peer \OR Prov
\Define PPTransit   = \Enter(PP) \AND \Exit(PP)
\Define CustTransit = \Later(Cust) \AND \Later(Prov)
\Define NoTransit   = \{
    \True \Path \NOT{}PPTransit \AND \NOT{}CustTransit
\}
\end{code}
\noindent
The \CD{NoTransit} constraint above ensures that requirements (P2) and (P5) are satisfied. In particular, it says that for any prefix, traffic should never both enter and exit the network from \CD{Peer} or \CD{Prov}. Similarly it prevents traffic from ever following paths that leave our network and later passing through both \CD{Prov} and \CD{Cust}.  To implement both \CD{Routing}
and \CD{NoTransit} simultaneously, we simply conjoin them:

%% \begin{lstlisting}[mathescape=true]
%% Routing $\AND$  NoTransit
%% \end{lstlisting}

\begin{code}
Routing \AND{} NoTransit
\end{code}

\noindent
The lines above capture the entire policy. From it, \sysname will generate per-device import and export filters, local preferences,
MED attributes, and community tags to ensure that the policy is
implemented correctly under all failure cases.

@@

In our data center,
there are three main concerns:
(1) traffic for the prefix block allocated to each top-of-rack router must be able to reach that router,
(2) local services must not leak outside the datacenter, and
(3) aggregation must be performed on global prefixes to reduce churn
in the network.

@@

%% \begin{lstlisting}[mathescape=true]
%% define Ownership = {
%%     $\path{p_{g1}}{end(A)}$
%%     $\path{p_{g2}}{end(B)}$
%%     $\path{p_{l1}}{end(E)}$
%%     $\path{p_{l2}}{end(F)}$
%% }
%% \end{lstlisting}

\begin{code}
\Define Ownership = \{
    PG1 \Path \End(A)
    PG2 \Path \End(B)
    PL1 \Path \End(E)
    PL2 \Path \End(F)
\}
\end{code}
\noindent
In English: traffic for prefix \CD{PG1} is only allowed to follow paths that
end at router \CD{A}; traffic not matching \CD{PG1}, 
but which matches \CD{PG2} must
end at router \CD{B}; and so on.

@@

%% \begin{lstlisting}[mathescape=true]
%% define Routing = {
%%     $\path{p_{g*}}{any}$
%%     $\path{p_{l*}}{\NOT enter(out)}$
%%     $\path{true}{exit(out)}$
%% }
%% \end{lstlisting}
\begin{code}
\Define Routing = \{
    PG1 \OR PG2 \Path \Any 
    PL1 \OR PL2 \Path \NOT{}\Enter(\Out)
    \True      \Path \Exit(\Out)
\}
\end{code}

\dpw{Note that I did not use an asterix above but used disjunction.}
The first line states that there is no restriction (\Any)
on how traffic must
traverse the network for global prefixes, aside from the default restriction
that traffic must not pass through the user's network and then loop
back on itself. This means traffic for
\CD{PG*} may be sent either from other routers in the datacenter, or
from external ASes. The second line ensures that traffic for local
prefixes never enters the network from an outside location. This constraint
guarantees that the services remain reachable only to locations
internal to the data center.

As in the backbone example, we can combine these constraints
logically to specify the network-wide policy.
However, in addition to constraints on the shape of paths,
\sysname allows the operator to specify constraints on the BGP control plane.
For instance, a constraint on aggregation is included to ensure that
aggregation for the prefixes in the set \CD{AGG} is performed from 
locations inside the network
to locations outside the network. \todo{Say more about aggregation here.
What prefixes are aggregated in to what blocks?  How is this defined?}

\begin{code}
Ownership \AND{} Routing \AND{} \Agg(AGG, \In \Link \Out)
\end{code}
